---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(modelr)
options(na.action = na.warn)
```

###Why are low quality diamonds more expensive?
```{r}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```
```{r}
ggplot(diamonds, aes(color, price)) + geom_boxplot()
```
```{r}
ggplot(diamonds, aes(clarity, price)) + geom_boxplot()
```
It looks like lower quality diamonds have higher prices because there is an important confounding variable: the weight (carat) of the diamond.
```{r}
ggplot(diamonds, aes(carat, price)) + 
  geom_hex(bins = 50)
```
We can make it easier to see how the other attributes of a diamond affect its relative price by fitting a model to separate out the effect of carat. But first, lets make a couple of tweaks to the diamonds dataset to make it easier to work with:

Focus on diamonds smaller than 2.5 carats (99.7% of the data)
Log-transform the carat and price variables.
```{r}
diamonds2 <- diamonds %>% 
  filter(carat <= 2.5) %>% 
  mutate(lprice = log2(price), lcarat = log2(carat))
ggplot(diamonds2, aes(lcarat, lprice)) + 
  geom_hex(bins = 50)
```
 Make the pattern explicit by fitting a model:
```{r}
mod_diamond <- lm(lprice ~ lcarat, data = diamonds2)
```

```{r}
grid <- diamonds2 %>% 
  data_grid(carat = seq_range(carat, 20)) %>% 
  mutate(lcarat = log2(carat)) %>% 
  add_predictions(mod_diamond, "lprice") %>% 
  mutate(price = 2 ^ lprice)

ggplot(diamonds2, aes(carat, price)) + 
  geom_hex(bins = 50) + 
  geom_line(data = grid, colour = "red", size = 1)
```
Now we can look at the residuals, which verifies that we’ve successfully removed the strong linear pattern:
```{r}
diamonds2 <- diamonds2 %>% 
  add_residuals(mod_diamond, "lresid")

ggplot(diamonds2, aes(lcarat, lresid)) + 
  geom_hex(bins = 50)
```
Importantly, we can now re-do our motivating plots using those residuals instead of price
```{r}
ggplot(diamonds2, aes(cut, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(color, lresid)) + geom_boxplot()
ggplot(diamonds2, aes(clarity, lresid)) + geom_boxplot()
```
###A more complicated model
```{r}
mod_diamond2 <- lm(lprice ~ lcarat + color + cut + clarity, data = diamonds2)
```

```{r}
grid <- diamonds2 %>% 
  data_grid(cut, .model = mod_diamond2) %>% 
  add_predictions(mod_diamond2)
grid

ggplot(grid, aes(cut, pred)) + 
  geom_point()
```

