---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(modelr)
options(na.action = na.warn)

library(nycflights13)
library(lubridate)
```
Let’s work through a similar process for a dataset that seems even simpler at first glance: the number of flights that leave NYC per day
```{r}
daily <- flights %>% 
  mutate(date = make_date(year, month, day)) %>% 
  group_by(date) %>% 
  summarise(n = n())
daily
```
```{r}
ggplot(daily, aes(date, n)) + 
  geom_line()
```
Let’s start by looking at the distribution of flight numbers by day-of-week:
```{r}
daily <- daily %>% 
  mutate(wday = wday(date, label = TRUE))
ggplot(daily, aes(wday, n)) + 
  geom_boxplot()
```
One way to remove this strong pattern is to use a model. First, we fit the model, and display its predictions overlaid on the original data:
```{r}
mod <- lm(n ~ wday, data = daily)

grid <- daily %>% 
  data_grid(wday) %>% 
  add_predictions(mod, "n")
```

```{r}
ggplot(daily, aes(wday, n)) + 
  geom_boxplot() +
  geom_point(data = grid, colour = "red", size = 4)
```
Next we compute and visualise the residuals:
```{r}
daily <- daily %>% 
  add_residuals(mod)

daily %>% 
  ggplot(aes(date, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line()
```
Our model seems to fail starting in June: you can still see a strong regular pattern that our model hasn’t captured. Drawing a plot with one line for each day of the week makes the cause easier to see:
```{r}
ggplot(daily, aes(date, resid, colour = wday)) + 
  geom_ref_line(h = 0) + 
  geom_line()
```
There are some days with far fewer flights than expected:
```{r}
daily %>% 
  filter(resid < -100)
```
f you’re familiar with American public holidays, you might spot New Year’s day, July 4th, Thanksgiving and Christmas. There are some others that don’t seem to correspond to public holidays. You’ll work on those in one of the exercises.
```{r}
daily %>% 
  ggplot(aes(date, resid)) + 
  geom_ref_line(h = 0) + 
  geom_line(colour = "grey50") + 
  geom_smooth(se = FALSE, span = 0.20)
```
There are fewer flights in January (and December), and more in summer (May-Sep). We can’t do much with this pattern quantitatively, because we only have a single year of data.

###Seasonal Saturday effect   
Let’s first tackle our failure to accurately predict the number of flights on Saturday. A good place to start is to go back to the raw numbers, focussing on Saturdays:
```{r}
daily %>% 
  filter(wday == "Sat") %>% 
  ggplot(aes(date, n)) + 
    geom_point() + 
    geom_line() +
    scale_x_date(NULL, date_breaks = "1 month", date_labels = "%b")
```
Lets create a “term” variable that roughly captures the three school terms, and check our work with a plot:
```{r}
term <- function(date) {
  cut(date, 
    breaks = ymd(20130101, 20130605, 20130825, 20140101),
    labels = c("spring", "summer", "fall") 
  )
}

daily <- daily %>% 
  mutate(term = term(date)) 

daily %>% 
  filter(wday == "Sat") %>% 
  ggplot(aes(date, n, colour = term)) +
  geom_point(alpha = 1/3) + 
  geom_line() +
  scale_x_date(NULL, date_breaks = "1 month", date_labels = "%b")
```
It’s useful to see how this new variable affects the other days of the week:
```{r}
daily %>% 
  ggplot(aes(wday, n, colour = term)) +
    geom_boxplot()
```
It looks like there is significant variation across the terms, so fitting a separate day of week effect for each term is reasonable. This improves our model, but not as much as we might hope:
```{r}
mod1 <- lm(n ~ wday, data = daily)
mod2 <- lm(n ~ wday * term, data = daily)

daily %>% 
  gather_residuals(without_term = mod1, with_term = mod2) %>% 
  ggplot(aes(date, resid, colour = model)) +
    geom_line(alpha = 0.75)
```
We can see the problem by overlaying the predictions from the model on to the raw data:
```{r}
grid <- daily %>% 
  data_grid(wday, term) %>% 
  add_predictions(mod2, "n")

ggplot(daily, aes(wday, n)) +
  geom_boxplot() + 
  geom_point(data = grid, colour = "red") + 
  facet_wrap(~ term)
```
Our model is finding the mean effect, but we have a lot of big outliers, so mean tends to be far away from the typical value. We can alleviate this problem by using a model that is robust to the effect of outliers: MASS::rlm(). This greatly reduces the impact of the outliers on our estimates, and gives a model that does a good job of removing the day of week pattern:
```{r}
mod3 <- MASS::rlm(n ~ wday * term, data = daily)

daily %>% 
  add_residuals(mod3, "resid") %>% 
  ggplot(aes(date, resid)) + 
  geom_hline(yintercept = 0, size = 2, colour = "white") + 
  geom_line()
```

